Make any stupid plot you want

### 2021-05-10 draft/sections/implementation.tex
closing out cite/spacing cleanups

\begin{minted}{python}
  fig, ax = plt.subplots()
  artist = ArtistClass(E, V)
  ax.add_artist(artist)
\end{minted}
Building on the current Matplotlib Artists which construct an internal representation of the graphic, \mintinline{python}{ArtistClass} acts as an equivalence class artist \vartisteq\ as described in \autoref{eq:math:artist:aprime}. The visual bundle \vtotal\ is specified as the \mintinline{python}{V} dictionary of the form \mintinline{python}|{parameter:(variable name, encoder)}| where parameter is a component in \vfiber, variable is a component in \dfiber, and the \vchannel\ encoders are passed in as functions or callable objects. The data bundle \dtotal\ is passed in as a \mintinline{python}{E} object. By binding data and transforms to \vartisteq\ inside \mintinline{python}{__init__}, the \mintinline{python}{draw} method is a fully specified artist \vartist\ as defined in \autoref{eq:math:artist:artist}.
\begin{minted}{python}
  class ArtistClass(matplotlib.artist.Artist): #A'
      def __init__(self, E, V, *args, **kwargs):
          # properties that are specific to the graphic
          self.E = E
          self.V = V 
          super().__init__(*args, **kwargs)

@@

      def draw(self, renderer):
          # returns K, indexed on fiber then key 
          tau = self.E.view(self.axes) 
          # visual channel encoding applied component wise 
          mu = {p: nu(tau(c))
                   for p, (c, nu) in self.V.items()} 
          self.q_hat(**mu)
          # pass configurations off to the renderer
          super().draw(renderer)
  \end{minted}
 The data is fetched in section \dsection\ via a \mintinline{python}{view} method on the data because the input to the artist is a section on \dtotal. The \mintinline{python}{view} method takes the \mintinline{python}{axes} attribute because it provides the region in graphic coordinates \gbase\ that can be used to query back into data to select a subset as described in \autoref{sec:math:data:sheaf}. We require that the \mintinline{python}{view} method be atomic so that we do not risk race conditions. Atomicity means that the values cannot change after the method is called in draw until a new call to draw\cite{ullmanFirstCourseDatabase2008}, which ensures the integrity of the section.

@@

 \subsection{Scatter, Line, and Bar Artists}
 \label{sec:code:artists}
 \begin{figure}[H]
    \centering 
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
        \caption{}
        \label{fig:code:scatter}
    \end{subfigure}
    \begin{subfigure}{.48\textwidth}
        \includegraphics[width=\textwidth]{figures/code/line_1.png}
        \caption{}
        \label{fig:code:line}
    \end{subfigure}
  \caption{Scatter plot and line plot implemented using \mintinline{python}{Point} and \mintinline{python}{Line} artists and fiber bundle inspired data models. Matplotlib is used for the rendering.}
  \label{fig:code:scatter_line}
\end{figure}
The figure in \autoref{fig:code:scatter} is described by \autoref{eq:eq:math:artist:q:scatter}. This is implemented via a \mintinline{python}{Line} object where the scatter marker shape is fixed as a circle, and the visual fiber components are x and y position and the facecolor and size of the marker. We only show the \mintinline{python}{q_hat} function here because the \mintinline{python}{__init__, draw} are inherited from the prototype artist \mintinline{Python}{ArtistClass}.

@@

\begin{minted}{python}
#\nu_i(m_r(E_i)) = \varphi(m_r)(\nu_i((E_i))
def test_nominal(values, encoder):
    m1 = list(zip(values, encoder(values)))
    random.shuffle(values)
    m2 = list(zip(values, encoder(values)))
    assert sorted(m1) == sorted(m2)
\end{minted}
As described in \autoref{eq:math:artist:nu:equivariance}, a test for equivariance can be implemented trivially. It is currently factored out of the artist for clarity.

\subsection{Data Model}
\label{sec:code:data}
The data input into the \mintinline{python}{Artist} will often be a wrapper class around an existing data structure. This wrapper object must specify the fiber components \dfiber\ and connectivity \dbase\ and have a \mintinline{python}{view} method that returns an atomic object that encapsulates \dsection. To support specifying the fiber bundle, we define a \mintinline{python}{FiberBundle} data class\cite{DataClasses}

@@

class LineData:
    def __init__(self):
        self.FB = FiberBundle({'tables': ['edge']},  
                     {'x': float, 'y': float, 'color': str})
    def tau(self, k):
        return # tau evaluated on interval k
\end{minted}
The discrete \mintinline{python}{tau} method returns a record of discrete points, akin to a row in a table, while the line\mintinline{python}{tau} returns a sampling of points along an edge \dbasepoint. These continuities are also specified in the \mintinline{python}|k={'tables':[]}| dictionary. 
\begin{minted}{python}
  def view(self, axes):
      table = defaultdict(list)
      for k in self.keys:
          table['index'].append(k)
          for (name, value) in zip(self.FB.fiber.keys(), 
                                   self.tau(k)[1]):
              table[name].append(value)
      return table
\end{minted}
In both cases, the \mintinline{python}{view} method packages the data into a data structure that the artist can unpack via data component name, akin to a table with column names when \dbase\ is 0 or 1 D.

@@

\subsection{Case Study: Penguins}
\label{sec:code_case_study}
Building on the \mintinline{python}{Bar} artist in \autoref{sec:code:artists}, we implement grouped bar charts as these do not exist out of the box in the current version of Matplotlib. Instead, grouped bar charts are often achieved via looping over an implementation of bar, which forces the user to keep track which values are mapped to a single visual element and how that is achieved. For this case study, we use the Palmer Penguins dataset\cite{gormanEcologicalSexualDimorphism2014, horstPalmerpenguinsPalmerArchipelago2020}, packaged as a pandas data frame\cite{nakhaeeMcnakhaeePalmerpenguins2021} since that is a very commonly used Python labeled data structure.

\begin{table}[H]
    \centering
\begin{tabular}{|l|r|r|r|l|l|l|}
    \toprule
       sex &  Adelie &  Chinstrap &  Gentoo & Adelie\_c & Chinstrap\_c & Gentoo\_c \\
    \midrule
     female &      73 &         34 &      58 &   Adelie &   Chinstrap &   Gentoo \\
     male &      73 &         34 &      61 &   Adelie &   Chinstrap &   Gentoo \\
    \bottomrule
\end{tabular}
\caption{Palmer Penguins dataset that is processed to become input into the grouped bar chart. This data is a count of penguin species. The columns with suffix c are used to specify the color of the corresponding visual element.}
\label{tab:code:penguins}
\end{table}

The wrapper is very thin because there is explicitly only one section.
\begin{minted}{python}
class DataFrame:
    def __init__(self, dataframe):
        self.FB = FiberBundle(K = {'tables':['vertex']},
                              F = dict(dataframe.dtypes))
        self._tau = dataframe.iloc
        self._view = dataframe

@@

Since all the visual transformation is passed through to \mintinline{python}{Bar}, the \mintinline{python}{draw} method does not do any visual transformations. In \mintinline{python}{StackedBar} the \mintinline{python}{view} is used to adjust the \mintinline{python}{floor} for every subsequent bar chart, since a stacked bar chart is bar chart area marks concatenated together in the length parameter. In contrast, \mintinline{python}{GroupedBar} does not even need the view, but instead keeps track of the relative position of each group of bars in the visual only variable \mintinline{python}{offset}.

>>>>

\begin{minted}{python}
  fig, ax = plt.subplots()
  artist = ArtistClass(E, V)
  ax.add_artist(artist)
\end{minted}
Building on the current Matplotlib Artists, which construct an internal representation of the graphic, \mintinline{python}{ArtistClass} acts as an equivalence class artist \vartisteq\ as described in \autoref{eq:math:artist:aprime}. The visual bundle \vtotal\ is specified as the \mintinline{python}{V} dictionary of the form \mintinline{python}|{parameter:(variable name, encoder)}| where parameter is a component in \vfiber, variable is a component in \dfiber, and the \vchannel\ encoders are passed in as functions or callable objects. The data bundle \dtotal\ is passed in as a \mintinline{python}{E} object. By binding data and transforms to \vartisteq\ inside \mintinline{python}{__init__}, the \mintinline{python}{draw} method is a fully specified artist \vartist\ as defined in \autoref{eq:math:artist:artist}.
\begin{minted}{python}
  class ArtistClass(matplotlib.artist.Artist): #A'
      def __init__(self, E, V, *args, **kwargs):
          # properties that are specific to the graphic
          self.E = E
          self.V = V 
          super().__init__(*args, **kwargs)

@@

      def draw(self, renderer):
          # returns K, indexed on fiber then key 
          tau = self.E.view(self.axes) 
          # visual channel encoding applied component wise 
          mu = {p: nu(tau(c))
                   for p, (c, nu) in self.V.items()} 
          self.q_hat(**mu)
          # pass configurations off to the renderer
          super().draw(renderer)
  \end{minted}
 The data in section \dsection\ is fetched via a \mintinline{python}{view} method on the data because the input to the artist is a section on \dtotal. The \mintinline{python}{view} method takes the \mintinline{python}{axes} attribute because it provides the region in graphic coordinates \gbase\ that can be used to query back into data to select a subset as described in \autoref{sec:math:data:sheaf}. We require that the \mintinline{python}{view} method be atomic so that we do not risk race conditions. Atomicity means that the values cannot change after the method is called in draw until a new call to draw\cite{ullmanFirstCourseDatabase2008}, which ensures the integrity of the section.

@@

 \subsection{Scatter, Line, and Bar Artists}
 \label{sec:code:artists}
 \begin{figure}[H]
    \centering 
    \begin{subfigure}{.49\textwidth}
        \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
        \caption{}
        \label{fig:code:scatter}
    \end{subfigure}
    \begin{subfigure}{.48\textwidth}
        \includegraphics[width=\textwidth]{figures/code/line_1.png}
        \caption{}
        \label{fig:code:line}
    \end{subfigure}
  \caption{Scatter plot and line plot implemented using \mintinline{python}{Point} and \mintinline{python}{Line} artists and fiber bundle inspired data models. Matplotlib is used for the rendering.}
  \label{fig:code:scatter_line}
\end{figure}
The figure in \autoref{fig:code:scatter} is a scatter plot, as described by \autoref{eq:eq:math:artist:q:scatter}. This is implemented via a \mintinline{python}{Line} object where the scatter marker shape is fixed as a circle, and the visual fiber components are x and y position and the facecolor and size of the marker. We only show the \mintinline{python}{q_hat} function here because the \mintinline{python}{__init__, draw} are inherited from the prototype artist \mintinline{Python}{ArtistClass}.

@@

\begin{minted}{python}
#\nu_i(m_r(E_i)) = \varphi(m_r)(\nu_i((E_i))
def test_nominal(values, encoder):
    m1 = list(zip(values, encoder(values)))
    random.shuffle(values)
    m2 = list(zip(values, encoder(values)))
    assert sorted(m1) == sorted(m2)
\end{minted}
As described in \autoref{eq:math:artist:nu:equivariance}, a test for equivariance can be implemented trivially. The test shown here is the nominal test described in \autoref{tab:math:artist:nu} because \mintinline{python}{Categorical} is intended to encode nominal data. Equivariance tests are currently factored out of the artist for clarity.

\subsection{Data Model}
\label{sec:code:data}
The data input into the \mintinline{python}{Artist} will often be a wrapper class around an existing data structure. This wrapper object must specify the fiber components \dfiber\ and connectivity \dbase\ and have a \mintinline{python}{view} method that returns an atomic object that encapsulates \dsection. To support specifying the fiber bundle, we define a \mintinline{python}{FiberBundle} data class \cite{DataClasses}

@@

class LineData:
    def __init__(self):
        self.FB = FiberBundle({'tables': ['edge']},  
                     {'x': float, 'y': float, 'color': str})
    def tau(self, k):
        return # tau evaluated on interval k
\end{minted}
The discrete \mintinline{python}{tau} method returns a record of discrete points, akin to a row in a table, while the line \mintinline{python}{tau} returns a sampling of points along an edge \dbasepoint. These continuities are also specified in the \mintinline{python}|k={'tables':[]}| dictionary. 
\begin{minted}{python}
  def view(self, axes):
      table = defaultdict(list)
      for k in self.keys:
          table['index'].append(k)
          for (name, value) in zip(self.FB.fiber.keys(), 
                                   self.tau(k)[1]):
              table[name].append(value)
      return table
\end{minted}
In both cases, the \mintinline{python}{view} method packages the data into a data structure that the artist can unpack via data component name, akin to a table with column names when \dbase\ is 0 or 1 D.

@@

\subsection{Case Study: Penguins}
\label{sec:code_case_study}
Building on the \mintinline{python}{Bar} artist in \autoref{sec:code:artists}, we implement grouped bar charts as these do not exist out of the box in the current version of Matplotlib. Instead, grouped bar charts are often achieved via looping over an implementation of bar, which forces the user to keep track which values are mapped to a single visual element and how that is achieved. For this case study, we use the Palmer Penguins dataset \cite{gormanEcologicalSexualDimorphism2014, horstPalmerpenguinsPalmerArchipelago2020}, packaged as a pandas data frame \cite{nakhaeeMcnakhaeePalmerpenguins2021} since that is a very commonly used Python labeled data structure.

\begin{table}[H]
    \centering
\begin{tabular}{|l|r|r|r|l|l|l|}
    \toprule
       sex &  Adelie &  Chinstrap &  Gentoo & Adelie\_c & Chinstrap\_c & Gentoo\_c \\
    \midrule
     female &      73 &         34 &      58 &   Adelie &   Chinstrap &   Gentoo \\
     male &      73 &         34 &      61 &   Adelie &   Chinstrap &   Gentoo \\
    \bottomrule
\end{tabular}
\caption{Palmer Penguins dataset that is processed to become input into the grouped bar chart. This data is a count of penguin species. The columns with suffix c are used to specify the color of the corresponding visual element.}
\label{tab:code:penguins}
\end{table}
The wrapper is very thin because there is explicitly only one section, which is the penguins dataframe. 
\begin{minted}{python}
class DataFrame:
    def __init__(self, dataframe):
        self.FB = FiberBundle(K = {'tables':['vertex']},
                              F = dict(dataframe.dtypes))
        self._tau = dataframe.iloc
        self._view = dataframe

@@

The visual transformations are passed through to \mintinline{python}{Bar}; therefore, the \mintinline{python}{draw} method does not do any visual transformations. In \mintinline{python}{StackedBar} the \mintinline{python}{view} is used to adjust the \mintinline{python}{floor} for every subsequent bar chart, since a stacked bar chart is bar chart area marks concatenated together in the length parameter. In contrast, \mintinline{python}{GroupedBar} does not even need the view, but instead keeps track of the relative position of each group of bars in the visual only variable \mintinline{python}{offset}.
